\documentclass{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{caption}
\usepackage{listings}
\usepackage{pdfpages}
\usepackage{amsmath,amssymb}

\newcommand{\qed}{\hfill $\blacksquare$}

\lstset{frame=single,keepspaces=true,captionpos=b}

\title{Homework 06 - Optimization}
\author{Arne Sachtler - \textit{Registration Number: 03692662}}
\date{\today}
\subtitle{IN2064 Machine Learning}

\begin{document}
\maketitle
	
\section{Convexity}
\subsection{Problem 1}
\begin{itemize}
	\item \textbf{i)}\\
	Given the function
	\begin{equation}
		f(x,y,z) = 3x + e^{y+z} - \min\{-x^2, \log(y)\} \text{ on } D = (-100,100)\times(1,50)\times(10,20)
	\end{equation}
	$f$ can be decomposed into a sum of three functions:
	\begin{equation}
		f(x,y,z) = g(x) + h(y,z) + k(x,y)\, ,
	\end{equation}
	where 
	\begin{enumerate}
		\item $g(x) = 3x$
		\item $h(y,z) = e^{y+z}$
		\item $k(x,y) = -\min(-x^2, \log(y))$ \, .
	\end{enumerate}
	The sum of functions is a convex function if the summands are convex. Here, $g$ is a linear function and linear functions are both convex and concave. The exponential map $h$ is convex, too. The interesting part is $k$, if it can be proven that $k$ is convex it can be concluded that $f$ is convex. Using the equation $-\min(x,y) = \max(-x,-y)$ the function $k$ can be rewritten to
	\begin{equation}
		k(x,y) = -\min(-x^2, \log(y)) = \max(x^2, -\log(y)) \, .
	\end{equation}
	The maximum of convex functions is also convex. We can follow that $k$ is convex as $x^2$ is convex on $(-100,100)$ (and everywhere else) and $-log(y)$ is convex on $(1,50)$. Therefore, $f$ is convex. \qed
	\item \textbf{ii)}\\
	Given the function
	\begin{equation}
		f(x,y) = yx^3 - 2yx^2 + y + 4 \text{ on } D = (-10, 10)^2
	\end{equation}
	we can compute the Hessian matrix in order to disprove convexity. First the gradient:
	\begin{equation}
		\nabla f(x,y) = \begin{pmatrix}
			3x^2y - 2xy&x^3-2x^2+1
		\end{pmatrix}\, ,
	\end{equation}
	then the Hessian is
	\begin{equation}
		\mathbf{H} = \begin{pmatrix}
			6xy-2y&3x^2-2x\\3x^2-2x&0
		\end{pmatrix} \, .
	\end{equation}
	A function $f$ is only convex if its Hessian is positive semi-definite. A matrix is positive semi-definite iff the eigenvalues are greater than or equal to zero. In order to disprove convexity we compute the eigenvalues of the Hessian using the characteristic polynomial
	\begin{equation}
		\det (\mathbf{H} - \lambda \mathbf{I}) = ... = \lambda^2 + (2y - 6xy)\lambda - (9x^4 - 12x^3 + 4x^2) \, .
	\end{equation}
	For the eigenvalues we get
	\begin{equation}
		\lambda_{1,2} = -\frac{2y-6xy}{2} \pm \sqrt{\frac{(2y-6xy)^2}{4} + 9x^4 - 12x^3 + 4x^2} \, .
	\end{equation}
	Consider the point $(0,1)\in D$. For the eigenvalues of the Hessian we get $\lambda_1 = 0$ and $\lambda_2 = -2 < 0$. It is proven that $f$ is not convex as $\mathbf{H}$ is not positive semi-definite. \qed
	\item \textbf{iii)}\\
	Consider
	\begin{equation}
		f(x) = \log(x) + x^3 \text{ on } D=(1, \infty) \, .
	\end{equation}
	To prove convexity, the second derivative can be computed:
	\begin{equation}
		f'(x) = \frac{1}{x} + 3x^2
	\end{equation}
	and
	\begin{equation}
		f''(x) = 6x - \frac{1}{x^2} \, .
	\end{equation}
	As it holds that
	\begin{equation}
		\forall x \in (1, \infty): 6x - \frac{1}{x^2} > 0
	\end{equation}
	we conclude that $f$ is convex. \qed
	\item \textbf{iv)}
	Similar to the first problem, the given function
	\begin{equation}
		f(x) = -\min(2\log(2x), -x^2+4x-32) \text{ on } D = \mathbb{R}^+
	\end{equation}
	can be rewritten in terms of the maximum function
	\begin{equation}
		f(x) = \max(-2\log(2x), x^2-4x+32) \, .
	\end{equation}
	Again, the maximum of two functions is convex if all functions that are considered for the maximum are convex.
	The function $g(x) = x^2 -4x+32$ is convex as $g''(x) = 2 > 0$. For $h(x) = -2\log(2x)$ we get $h'(x) = -\frac{2}{x}$ and $h''(x) = \frac{2}{x^2} \ge 0 \forall x \in \mathbb{R}^+$. Therefore, $f$ is convex. \qed
\end{itemize}
\end{document}
